DeepPicar: A Low-cost Deep Neural Network-based Autonomous Car

### 2018-04-04 paper/evaluation.tex
Minor changes
Initial commit

The Raspberry Pi 3 Model B platform used in DeepPicar equips a Boardcom
BCM2837 SoC, which has a quad-core ARM Cortex-A53 cluster,
running at up to 1.2GHz. Each core has a 32K private Instruction cache
and a 32K private Data cache, and all cores share a 512KB L2 cache.
The chip also includes Broadcom's Videocore IV
GPU, although we did not use the GPU in our evaluation due to the lack
of sofware support (TensorFlow is not compatible with the Raspberry Pi's GPU).
For software, we use Ubuntu MATE 16.04, TensorFlow 1.1 and Python
2.7. We disabled DVFS (dynamic voltage frequency scaling) and
configured the clock speed of each core statically at the maximum 1.2GHz.

@@

For real-time control of a car (or any robot), the control loop
frequency must be sufficiently high so that the car can quickly
react to the changing environment and its internal states. In general,
control performance improves when the frequency is higher, though
computation time and the type of the particular physical system are
factors in determining a proper control loop frequency. While a standard
control system may be comprised of multiple control loops with
differing control frequencies---e.g., an inner control loop for lower-level
PD control, an outer loop for motion planning, etc.---DeepPicar's
control loop is a single layer, as shown earlier in
Figure~\ref{fig:controlloop}, since a single deep neural network
replaces the traditional multi-layer control pipline. (Refer to
Figure~\ref{fig:end-to-end-control} on the differences between the
standard robotics control vs. end-to-end deep learning approach).
This means that the DNN inference operation must be completed
within the inner-most control loop update frequency. To understand
achievable control-loop update frequencies, we experimentally measured
the execution times of DeepPicar's DNN inference operations.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=.75\textwidth]{figs/Fig7_new}
  \caption{DeepPicar's control loop processing times over 1000 input image frames.}
  \label{fig:control-loop-timing}
\end{figure}

@@

Second, we find that the measured average
execution time of a single control loop is 22.87 ms, or 43.7 Hz and
the 99 percentile time is 23.38 ms.
This means that the DeepPicar can operate
at a up to 40 Hz control frequency in real-time using only the on-board
Raspberry Pi 3 computing platform, as no remote computing resources 
were necessary. We consider these results surprising given the complexity
of the deep neural network, and the fact that the inference operation
performed by TensorFlow only utilizes the CPU cores of the
Raspberry Pi 3 (its GPU is not supported by Tensorflow).
In comparison, NVIDIA's DAVE-2 system, which has the exact same neural
network architecture, reportedly runs at 30 Hz~\cite{Bojarski2016}. 
Although we believe it was not
limited by their computing platform (we will experimentally compare
performance differences among multiple embedded computing platforms,
including NVIDIA's Jetson TX2, later in
Section~\ref{sec:comparison}), the fact that a low-cost
Raspberry Pi 3 can achieve comparable real-time control performance is
surprising.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=.7\textwidth]{figs/perf_vs_corecnt}
  \caption{Average control loop execution time vs. \#of CPU
    cores.}
  \label{fig:perf-vs-corecnt}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=.7\textwidth]{figs/perf_vs_modelcnt}
  \caption{Timing impact of co-scheduling multiple DNNs. 1Nx1C: one DNN
    model using one core; 4Nx1C: four DNN models each using one core;
    1Nx2C: one DNN model using two cores; 2Nx2C: two DNN models each
    using two cores.} 
  \label{fig:perf-vs-modelcnt}
\end{figure}

@@

In this experiment, we investigate the impact of contended DRAM
requests to the DNN inference timing of DeepPicar.
For the experiment, we use a synthetic memory
intensive benchmark from the IsolBench suite~\cite{Valsan2016}.
We run a single DNN model on one core, and
co-schedule an increasing number of the memory intensive synthetic
benchmarks~\footnote{We use the \emph{Bandwidth} benchmark in the
  IsolBench suite, with the following command line parameters: \texttt{\$
  bandwdith -a write -m 16384}}, on the remaining idle cores.

\begin{figure}[h]
  \centering
  \includegraphics[width=.7\textwidth]{figs/perf_vs_bandwidth}
  \caption{Effect of memory performance hogs on the DNN
    inferencing. The DNN model uses Core 0 and memory-hog co-runners
    use the rest of the cores.}
  \label{fig:}
\end{figure}

Figure~\ref{fig:} shows the normalized execution time (first group) and
L2 miss-rate (last group)
of the DNN model running on the Core 0 as a function of the number of
co-scheduled memory intensive synthetic benchmarks. We also shows the
normalized total system bus access count (middle group), which is
representative of memory bandwidth usage of all the cores.
First, as we
increase the number of co-runners, the DNN model's execution times are
increased exponentially---by up to 11.6X---even though the DNN model
is running on a dedicated core (Core 0). On the other hand, the DNN
model's L2 cache-miss rates do not increase as much.
In other words, the DNN model's execution increase is not fully
explained by increased contention in the L2 cache space.
Instead, the exponential increase in the total number of bus accesses,
due to the memory-hog co-runners, appears to be more closely
correlated with the DNN model's execution time increase. This is
the high memory bandwidth pressure from the co-scheduled
memory intensive benchmarks would have increased the DNN model's
memory access latency.

@@

%% \input{evaluation-cachepart}

@@

%% Lastly, we find that partitioning the shared L2 cache provides 
%% practically no benefits to the real-time performance of the DNN. To the 
%% contrary, performance is constant regardless of the amount of cache 
%% space used. We also find that performance becomes worse when memory
%% intensive co-runners are introduced as they generate more memory 
%% contention when partitioning is enabled. Furthermore, we 
%% find that performance remains consistent when multiple models are run 
%% simultaneously. As such, we conclude that the shared L2 cache is not 
%% essential to the DNN's real-time performance.

>>>>

The Raspberry Pi 3 Model B platform used in DeepPicar equips a Broadcom
BCM2837 SoC, which has a quad-core ARM Cortex-A53 cluster,
running at up to 1.2GHz. Each core has a 32K private Instruction cache
and a 32K private Data cache, and all cores share a 512KB L2 cache.
The chip also includes Broadcom's Videocore IV
GPU, although we did not use the GPU in our evaluation due to the lack
of sofware support (TensorFlow is not compatible with the Raspberry Pi's GPU).
For software, we use Ubuntu MATE 16.04, TensorFlow 1.1 and Python
2.7. We disabled DVFS (dynamic voltage frequency scaling) and
configured the clock speed of each core statically at the maximum 1.2GHz.

@@

For real-time control of a car (or any robot), the control loop
frequency must be sufficiently high so that the car can quickly
react to the changing environment and its internal states. In general,
control performance improves when the frequency is higher, though
computation time and the type of particular physical system are
factors in determining a proper control loop frequency. While a standard
control system may be comprised of multiple control loops with
differing control frequencies---e.g., an inner control loop for lower-level
PD control, an outer loop for motion planning, etc.---DeepPicar's
control loop is a single layer, as shown earlier in
Figure~\ref{fig:controlloop}, since a single deep neural network
replaces the traditional multi-layer control pipline. (Refer to
Figure~\ref{fig:end-to-end-control} on the differences between the
standard robotics control vs. end-to-end deep learning approach).
This means that the DNN inference operation must be completed
within the inner-most control loop update frequency. To understand
achievable control-loop update frequencies, we experimentally measured
the execution times of DeepPicar's DNN inference operations.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=.5\textwidth]{figs/Fig7_new}
  \caption{DeepPicar's control loop processing times over 1000 input image frames.}
  \label{fig:control-loop-timing}
\end{figure}

@@

Second, we find that the measured average
execution time of a single control loop is 22.87 ms, or 43.7 Hz and
the 99 percentile time is 23.38 ms.
This means that the DeepPicar can operate
at up to 40 Hz control frequency in real-time using only the on-board
Raspberry Pi 3 computing platform, as no remote computing resources 
were necessary. We consider these results surprising given the complexity
of the deep neural network, and the fact that the inference operation
performed by TensorFlow only utilizes the CPU cores of the
Raspberry Pi 3 (its GPU is not supported by Tensorflow).
In comparison, NVIDIA's DAVE-2 system, which has the exact same neural
network architecture, reportedly runs at 30 Hz~\cite{Bojarski2016}. 
Although we believe it was not
limited by their computing platform (we will experimentally compare
performance differences among multiple embedded computing platforms,
including NVIDIA's Jetson TX2, later in
Section~\ref{sec:comparison}), the fact that a low-cost
Raspberry Pi 3 can achieve comparable real-time control performance is
surprising.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/perf_vs_corecnt}
  \caption{Average control loop execution time vs. \#of CPU
    cores.}
  \label{fig:perf-vs-corecnt}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/perf_vs_modelcnt}
  \caption{Timing impact of co-scheduling multiple DNNs. 1Nx1C: one DNN
    model using one core; 4Nx1C: four DNN models each using one core;
    1Nx2C: one DNN model using two cores; 2Nx2C: two DNN models each
    using two cores.} 
  \label{fig:perf-vs-modelcnt}
\end{figure}

@@

In this experiment, we investigate the impact of contended DRAM
requests to the DNN inference timing of DeepPicar.
For the experiment, we use a synthetic memory
intensive benchmark from the IsolBench suite~\cite{Valsan2016}.
We run a single DNN model on one core, and
co-schedule an increasing number of the memory intensive synthetic
benchmarks~\footnote{We use the \emph{Bandwidth} benchmark in the
  IsolBench suite, with the following command line parameters: \texttt{\$
  bandwdith -a write -m 16384 -t 1000}}, on the remaining idle cores.

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/perf_vs_bandwidth}
  \caption{Average processing time vs. the number of memory
    memory intensive co-runners introduced.}
  \label{fig:perf_vs_bandwidth}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/perf_vs_bandwidth_mem}
  \caption{Effect of memory performance hogs on the shared resources. 
    The DNN model uses Core 0 and memory-hog co-runners
    use the rest of the cores.}
  \label{fig:perf_vs_bandwidth_mem}
\end{figure}

Figure~\ref{fig:perf_vs_bandwidth} shows the execution time of the DNN 
model and Figure~\ref{fig:perf_vs_bandwidth_mem} shows both the normalized
total memory bus accesses (first group) and L2 miss rates (second group).
In both figures, the DNN model is running on Core 0 as a function of the 
number of co-scheduled memory intensive synthetic benchmarks.
%% We also show 
%% the normalized total system bus access count (middle group), which is
%% representative of memory bandwidth usage of all the cores.
First, as we
increase the number of co-runners, the DNN model's execution times are
increased exponentially---by up to 11.6X---even though the DNN model
is running on a dedicated core (Core 0). On the other hand, the DNN
model's L2 cache-miss rates do not increase as much.
In other words, the DNN model's execution increase is not fully
explained by increased contention in the L2 cache space.
Instead, the exponential increase in the total number of bus accesses,
due to the memory-hog co-runners, appears to be more closely
correlated with the DNN model's execution time increase. This means
the high memory bandwidth pressure from the co-scheduled
memory intensive benchmarks increased the DNN model's
memory access latency.

@@

\input{evaluation-cachepart}
\input{evaluation-drampart}

@@

Fifth, we find that partitioning the shared L2 cache provides 
practically no benefits to the real-time performance of the DNN. To the 
contrary, performance is constant regardless of the amount of cache 
space used. Furthermore, we find that performance remains consistent 
when multiple models are run simultaneously, and when memory intensive
read co-runners are introduced. As such, we conclude that 
the shared L2 cache is not essential to the DNN's real-time performance.

Lastly, we find that memory bandwidth throttling results in significant
changes to model performance. When less bandwidth is given to models,
whether they're running alone or concurrently, we find that their performance
decreases noticably. When memory intensive co-runners are
present, we find that limiting the available bandwidth of the co-runners
actually results in improved model performance. As such, we conclude
that the shared DRAM memory is essential to the DNN's real-time 
performance.

### 2018-04-04 paper/intro.tex
Initial commit

% our goals
Instead, we want to build a low cost testbed that still employs the
state-of-the art AI technologies. Specifically, we focus on a end-to-end
deep learning based real-time control system,
which was developed for a real self-driving car, NVIDIA
DAVE-2~\cite{Bojarski2016}, and use the same methodology on a
smaller and less costly setup. In developing the testbed, our
goals are (1) to analyze real-time issues in DNN based end-to-end
control; and (2) to evaluate real-time performance of contemporary embedded
platforms for such workload.

% DeepPicar introduction
In this paper, we present DeepPicar, a low-cost autonomous car
platform for research. From a hardware perspective,
DeepPicar is comprised of a Raspberry Pi 3 Model B quad-core
computer, a web camera and a RC car, all of which are affordable
components (less than \$100 in total).
The DeepPicar, however, employs state-of-the-art AI
technologies, including a vision-based end-to-end control system that
utilizes a deep convolutional neural network (CNN).
The network receives an image frame from a single forward
looking camera as input and generates a predicted steering angle
value as output at each control period in \emph{real-time}.
The network has 9 layers, about 27 million connections
and 250 thousand parameters (weights).
The network architecture is almost identical to that of NVIDIA's DAVE-2
self-driving car~\cite{Bojarski2016}, which uses a much more powerful
computer (Drive PX computer~\cite{drivepx}) than a Raspberry Pi~3.
We chose to use a Pi 3 not only because it is affordable, but also because it is representative
of today's mainstream low-end embedded multicore platforms found in
smartphones and other embedded devices.

@@

% Findings:
%
% - we find DNN processing is an ideal real-time workload from the
%   perspective of timing predictability as the amount of computation
%   needed is fixed at design time and doesn't change over input.
%
% - real-time processing of non-trivial DNN inferencing is feasible
%   on today's embedded multicore platform, even as inexpensive as
%   raspberry pi3.
%
% - all tested platforms have enough computing capacity to consolidate 
%   multiple DeepPicar's CNN workloads or other compute intensive tasks
%   simultaneously.
%
% - however, when consolidating multiple workloads shared resource
%   contention is an important issue. we carefully analyzed
%   sensitivity of DeepPicar's workload with respect to shared
%   hardware resources.
%
% - in particular, we find DNN processing is highly sensitive to
%   memory performance. performance severely affected when memory
%   bandwidth is contented.
%
% - however, we find that dnn processing is not sensitive to cache
%   space. we find applying cache partitioning does not help improve
%   isolation of DNN workload when memory bandwidth is contended.
From the systematic study, we made a number of interesting
observations from the perspective of real-time systems.
% predictability
First, we find that DNN inferencing is highly predictable---from the
timing perspective---as the amount of computation needed to complete a
single inference is fixed at the DNN architecture design time and does
not change at runtime over different inputs (e.g., different image
frames). This predictable timing behavior is obviously a desirable
property for real-time systems.
%% , which gives another reason for the
%% real-time community to be interested in the use of DNNs.

% performance on low-cost embedded multicore
Second, we find that real-time processing of DeepPicar's CNN is
feasible on today's embedded computing platforms, even  
as inexpensive as the Raspberry Pi 3. We find that the control loop 
completes in under 33.$\overline{\mbox{33}}$ ms, or 30 Hz, using just two 
cores of the Raspberry Pi 3's quad-core CPU (w/o using its GPU), and can
do so 100\% of the time; or we can achieve 20 Hz control frequency using
just one core.
Other tested embedded platforms, the Intel UP and NVIDIA TX2, offer even
better performance, and are capable of supporting deep-learning based 
real-time control over 100 Hz control frequency on the TX2 when its
embedded GPU is used.
We find that all the tested embedded computing platforms have 
enough computing capacity to consolidate multiple instances of
DeepPicar's DNN workload.

% all platforms have enough capacity to consolidate
Third, contention in shared hardware resources remains an important
issue when consolidating multiple workloads. 
In particular, we find that DNN inferencing is highly sensitive to
memory performance interference as we observe up to an 11.6X control
loop execution time increase, when memory performance intensive
applications are co-scheduled on idle CPU cores.
On the other hand, we find that DNN inferencing is not sensitive to
cache space.
%% ; we show that applying cache partitioning has negligible impact
%% in providing timing isolation on consolidated scenarios.
This finding
suggests that in order to guarantee real-time performance of DNN-based
control applications, solutions to guarantee their memory performance
are more important.

% contributions
The {\bf contributions} of this paper are as follows:
\begin{itemize}
  \item We present DeepPicar, a low-cost autonomous car testbed, which
    is based on a state-of-the-art end-to-end deep learning, as 
    open-source~\footnote{The source code, and build 
instructions for the DeepPicar
can be found at: \url{https://github.com/heechul/picar}}, with the
    goal of lowering the economic and safety-related barriers to study
    autonomous cars.

  \item We systematically analyze real-time characteristics of DeepPicar's
    DNN inferencing workload on a number of representative
    contemporary embedded computing platforms. To the
    best of our knowledge, no prior work investigated the
    impacts of shared resource contention on DNN inferencing workloads
    in the context of real-time systems.

@@

The remaining sections of the paper are as follows.
Section~\ref{sec:background}
provides a background of applications in autonomous driving and related works.
Section~\ref{sec:overview} gives an overview of the DeepPicar
platform, including the high-level system and the methods used for
training and inference. Section~\ref{sec:evaluation} presents our
evaluation of the platform and how different factors can affect
performance. Section~\ref{sec:comparison} gives a comparison between  
the Raspberry Pi 3 and other embedded computing platforms to 
determine their suitably for autonomous driving research.
We discuss more detailed training methods and system-level issues
in~\ref{sec:discussion}. We review related work in
Section~\ref{sec:related} and conclude in
Section~\ref{sec:conclusion}.

>>>>

% our goals
Instead, we want to build a low cost testbed that still employs the
state-of-the art AI technologies. Specifically, we focus on an end-to-end
deep learning based real-time control system,
which was developed for a real self-driving car, NVIDIA
DAVE-2~\cite{Bojarski2016}, and uses the same methodology on a
smaller and less costly setup. In developing the testbed, our
goals are (1) to analyze real-time issues in DNN based end-to-end
control; and (2) to evaluate real-time performance of contemporary embedded
platforms for such a workload.

% DeepPicar introduction
In this paper, we present DeepPicar, a low-cost autonomous car
platform for research. From a hardware perspective,
DeepPicar is comprised of a Raspberry Pi 3 Model B quad-core
computer, a web camera and a RC car, all of which are affordable
components (less than \$100 in total).
The DeepPicar, however, employs state-of-the-art AI
technologies, including a vision-based end-to-end control system that
utilizes a deep convolutional neural network (CNN).
The network receives an image frame from a single forward
looking camera as input and generates a predicted steering angle
value as output at each control period in \emph{real-time}.
The network has 9 layers, about 27 million connections
and 250 thousand parameters (weights).
The network architecture is almost identical to that of NVIDIA's DAVE-2
self-driving car~\cite{Bojarski2016}, which uses a significantly more powerful
computer (Drive PX computer~\cite{drivepx}) than a Raspberry Pi~3.
We chose to use a Pi 3 not only because it is affordable, but also because it is representative
of today's mainstream low-end embedded multicore platforms found in
smartphones and other embedded devices.

@@

% Findings:
%
% - we find DNN processing is an ideal real-time workload from the
%   perspective of timing predictability as the amount of computation
%   needed is fixed at design time and doesn't change over input.
%
% - real-time processing of non-trivial DNN inferencing is feasible
%   on today's embedded multicore platform, even as inexpensive as
%   raspberry pi3.
%
% - all tested platforms have enough computing capacity to consolidate 
%   multiple DeepPicar's CNN workloads or other compute intensive tasks
%   simultaneously.
%
% - however, when consolidating multiple workloads shared resource
%   contention is an important issue. we carefully analyzed
%   sensitivity of DeepPicar's workload with respect to shared
%   hardware resources.
%
% - in particular, we find DNN processing is highly sensitive to
%   memory performance. performance severely affected when memory
%   bandwidth is contented.
%
% - however, we find that dnn processing is not sensitive to cache
%   space. we find applying cache partitioning does not help improve
%   isolation of DNN workload when memory bandwidth is contended.
From the systematic study, we have made a number of interesting
observations from the perspective of real-time systems.
% predictability
First, we find that DNN inferencing is highly predictable---from the
timing perspective---as the amount of computation needed to complete a
single inference is fixed at the DNN architecture design time and does
not change at runtime over different inputs (e.g., different image
frames). This predictable timing behavior is obviously a desirable
property for real-time systems.
%% , which gives another reason for the
%% real-time community to be interested in the use of DNNs.

% performance on low-cost embedded multicore
Second, we find that real-time processing of DeepPicar's CNN is
feasible on today's embedded computing platforms, even one  
as inexpensive as the Raspberry Pi 3. We find that the control loop 
completes in under 33.$\overline{\mbox{33}}$ ms, or 30 Hz, using just two 
cores of the Raspberry Pi 3's quad-core CPU (w/o using its GPU), and can
do so 100\% of the time; or we can achieve 20 Hz control frequency using
just one core.
Other tested embedded platforms, the Intel UP and NVIDIA TX2, offer even
better performance, and are capable of supporting deep-learning based 
real-time control over 100 Hz control frequency on the TX2 when its
embedded GPU is used.
We find that all the tested embedded computing platforms have 
enough computing capacity to consolidate multiple instances of
DeepPicar's DNN workload.

% all platforms have enough capacity to consolidate
Third, contention in shared hardware resources remains an important
issue when consolidating multiple workloads. 
In particular, we find that DNN inferencing is highly sensitive to
memory performance interference, as we observe up to an 11.6X control
loop execution time increase when memory performance intensive
applications are co-scheduled on idle CPU cores.
On the other hand, we find that DNN inferencing is not sensitive to
cache space.
%% ; we show that applying cache partitioning has negligible impact
%% in providing timing isolation on consolidated scenarios.
This finding
suggests that in order to guarantee real-time performance of DNN-based
control applications, solutions to guarantee their memory performance
are more important.

% contributions
The {\bf contributions} of this paper are as follows:
\begin{itemize}
  \item We present DeepPicar, a low-cost autonomous car testbed, which
    is based on a state-of-the-art end-to-end deep learning, as 
    open-source~\footnote{The source code, and build 
instructions for the DeepPicar
can be found at: \url{https://github.com/heechul/picar}}, with the
    goal of lowering the economic and safety-related barriers to the study
    of autonomous cars.

  \item We systematically analyze real-time characteristics of DeepPicar's
    DNN inferencing workload on a number of representative
    contemporary embedded computing platforms. While larger embedded
    platforms have been evaluated, as in  \cite{Otterness2017}, to the
    best of our knowledge, no prior work investigated the
    impacts of shared resource contention on DNN inferencing workloads
    in the context of smaller embedded platforms like the Raspberry Pi 3.

@@

The remaining sections of the paper are as follows.
Section~\ref{sec:background}
provides a background of applications in autonomous driving and related works.
Section~\ref{sec:overview} gives an overview of the DeepPicar
platform, including the high-level system and the methods used for
training and inference. Section~\ref{sec:evaluation} presents our
evaluation of the platform and how different factors can affect
performance. Section~\ref{sec:comparison} offers a comparison between  
the Raspberry Pi 3 and other embedded computing platforms to 
determine their suitably for autonomous driving research.
%% We discuss more detailed training methods and system-level issues
%% in~\ref{sec:discussion}. 
We review related work in
Section~\ref{sec:related} and conclude in
Section~\ref{sec:conclusion}.

### 2018-04-04 paper/overview.tex
Initial commit

\begin{figure}[h]
  \centering
  \includegraphics[width=.65\textwidth]{figs/DeepPicar_platform}
  \caption{DeepPicar platform.}
  \label{fig:overview}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.65\textwidth]{figs/architecture}
  \caption{DeepPicar's neural network architecture: 9 layers (5
    convolutional, 4 fully-connected layers), 27 million connections,
    250K parameters. The CNN architecture is identical to the one 
	used in NVIDIA's real self-driving car~\cite{Bojarski2016}.}
  \label{fig:architecture}
\end{figure}

For the neural network architecture, we implement
NVIDIA DAVE-2's convolutional neural network (CNN) using a TensorFlow
based CNN model in ~\cite{deeptesla}. Note, however, that the CNN model
in~\cite{deeptesla} was considerably bigger than NVIDIA's CNN model as
it contained an additional fully-connected layer of approximately 1.3M
additional parameters. We eliminated the additional layer to
faithfully recreate the NVIDIA's original CNN model.
%% The main difference is that we do not utilize a normalization layer, and 
%% instead initialize the weights using a Xavier initialization. 
As in DAVE-2, the CNN takes a raw color image (200x66 RGB pixels)
as input and produces a single steering angle value as
output. Figure~\ref{fig:architecture} shows the network architecture
used in this paper, which is comprised of 9 layers, 250K parameters,
and about 27 million connections as in NVIDIA DAVE-2's. Note, however,
that we did not apply the normalization mentioned
in~\cite{Bojarski2016}, as it does not include trainable parameters
and its computational demand with respect to the overall CNN
processing is minimal.

\begin{figure}[h]
  \centering
%% <<<<<<< HEAD
%%   \includegraphics[width=.5\textwidth]{figs/track_new2}
%%   \caption{One of the custom tracks used for training/testing.}
%%   \label{fig:track}
%% =======
  \begin{subfigure}{0.55\textwidth}
    \includegraphics[width=\textwidth]{figs/track_new2}
    \caption{Track 1.}
    \label{fig:track}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.7\textwidth}
    \includegraphics[width=\textwidth]{figs/track-2}
    \caption{Track 2.}
    \label{fig:track2}
  \end{subfigure}
  \caption{Custom tracks used for training/testing}
  \label{fig:tracks}
%% >>>>>>> f3814396d8701264f2439b42fbe8bbb608c41452
\end{figure}

% data collection and training.
To collect the training data, a human pilot manually drives the RC car
on tracks we created (Figure~\ref{fig:tracks}) to record
timestamped videos and contol commands. The stored data is then copied 
to a desktop computer, which is equipped with a NVIDIA GTX 1050 Ti GPU, 
where we train the network to accelerate training speed. 
For comparison, training the network on the Raspbeery Pi 3 takes
approximately 4 hours, whereas it takes only about 4 minutes on the
desktop computer using the GTX 1050 Ti GPU.

@@

% inferencing on pi3
Once the network is trained on the desktop computer, the trained model
is copied back to the Raspberry Pi 3. The network is then used
by the car's main controller, which feeds a image frame from the web
camera as input to the network. In each control period, the produced 
steering angle output is then converted as the PWM values of the 
steering motor of the car. Figure~\ref{fig:controlloop} shows simplified 
pseudo code of the controller's main loop. Among the five steps, the 3rd step, 
network inferencing, is the most computationally intensive and dominates the
execution time.

@@

Other issues we observed in training/testing the network, which
can affect the prediction accuracy of the CNN, are camera and actuator
(motor) control latencies. In DeepPicar's
context, the camera latency is from the time the camera
sensor observes the scene to the time the computer actually reads the
digitized image data. This time can be noticable depending on the camera 
used and the performance of the Pi. Higher camera latency could
negatively affect control performance, because the DNN would analyze
stale scenes. Actuator (motor) control latency also can be a
factor, because moving the steering motor to a desired position takes
time. 
In our platform, the combined latency is measured to be around
50 ms. In other words, in the recorded video and control data, a
control action (left, center, right) taken by the CNN appears to be
applied 50 ms later. Considering that camera's framerate is 30Hz
(33.$\overline{\mbox{33}}$ ms/frame), this is about two frames after the 
control action.
%% We experimentally measured the camera
%% latency and found it to be around 50-100 ms.
If this value is too high, control performance may suffer.
Our initial prototype suffered this problem as the observed camera and
control latency was as high as 300 ms, which negatively affected
control performance.
For reference, the latency of human perception is known to be as fast
as 13 ms~\cite{ThomasBurger2015}. 
% https://www.pubnub.com/blog/2015-02-09-how-fast-is-realtime-human-perception-and-technology/

Despite these issues, the trained models were still able to
achieve a reasonable degree of accuracy, successfully navigating
several different tracks we used for training. In our testing, we found 
that the DeepPicar could remain on a fairly challenging track for over 
10 minutes at a moderate speed (50\% throttle), at which point we stopped
the experiment, and more than one minute at a higher speed (75\%
throttle)~\footnote{A collection of self-driving videos can be found
  at: \url{https://photos.app.goo.gl/ce93sU7jPk4ywO8u2}}. Running at
higher speed is inherently more challenging because the CNN controller
has less time to recover from mistakes (bad predictions).  Also, we
find that the prediction accuracy is significantly affected by the
quality of training data as well as various environmental factors such
as the lighting condition. We plan to investigate more systematic ways
to improve the CNN's prediction accuracies.

>>>>

\begin{figure}[h]
  \centering
  \includegraphics[width=.4\textwidth]{figs/DeepPicar_platform}
  \caption{DeepPicar platform.}
  \label{fig:overview}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.4\textwidth]{figs/architecture}
  \caption{DeepPicar's neural network architecture: 9 layers (5
    convolutional, 4 fully-connected layers), 27 million connections,
    250K parameters. The CNN architecture is identical to the one 
	used in NVIDIA's real self-driving car~\cite{Bojarski2016}.}
  \label{fig:architecture}
\end{figure}

For the neural network architecture, we implement
NVIDIA DAVE-2's convolutional neural network (CNN) using a TensorFlow
based CNN model in ~\cite{deeptesla}. Note, however, that the CNN model
in~\cite{deeptesla} was considerably larger than NVIDIA's CNN model as
it contained an additional fully-connected layer of approximately 1.3M
additional parameters. We eliminated the additional layer to
faithfully recreate the NVIDIA's original CNN model.
%% The main difference is that we do not utilize a normalization layer, and 
%% instead initialize the weights using a Xavier initialization. 
As in DAVE-2, the CNN takes a raw color image (200x66 RGB pixels)
as input and produces a single steering angle value as
output. Figure~\ref{fig:architecture} shows the network architecture
used in this paper, which is comprised of 9 layers, 250K parameters,
and about 27 million connections as in NVIDIA DAVE-2's. Note, however,
that we did not apply the normalization mentioned
in~\cite{Bojarski2016}, as it does not include trainable parameters
and its computational demand with respect to the overall CNN
processing is minimal.

\begin{figure}[h]
  \centering
  \begin{subfigure}{0.4\textwidth}
    \includegraphics[width=\textwidth]{figs/track_new2}
    \caption{Track 1.}
    \label{fig:track}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.4\textwidth}
    \includegraphics[width=\textwidth]{figs/track-2}
    \caption{Track 2.}
    \label{fig:track2}
  \end{subfigure}
  \caption{Custom tracks used for training/testing}
  \label{fig:tracks}
\end{figure}

% data collection and training.
To collect the training data, a human pilot manually drives the RC car
on tracks we created (Figure~\ref{fig:tracks}) to record
timestamped videos and contol commands. The stored data is then copied 
to a desktop computer, which is equipped with a NVIDIA GTX 1060 GPU, 
where we train the network to accelerate training speed. 
For comparison, training the network on the Raspbeery Pi 3 takes
approximately 4 hours, whereas it takes only about 4 minutes on the
desktop computer using the GTX 1050 Ti GPU.

@@

% inferencing on pi3
Once the network is trained on the desktop computer, the trained model
is copied back to the Raspberry Pi 3. The network is then used
by the car's main controller, which feeds an image frame from the web
camera as input to the network. In each control period, the produced 
steering angle output is then converted into the PWM values of the 
steering motor of the car. Figure~\ref{fig:controlloop} shows simplified 
pseudo code of the controller's main loop. Among the five steps, the 3rd step, 
network inferencing, is the most computationally intensive and dominates the
execution time.

@@

Other issues we observed in training/testing the network, which
can affect the prediction accuracy of the CNN, are camera and actuator
(motor) control latencies. In DeepPicar's
context, the camera latency  measured from the time the camera
sensor observes the scene to the time the computer actually reads the
digitized image data. This time can be noticable depending on the camera 
used and the performance of the Pi. Higher camera latency could
negatively affect control performance, because the DNN would analyze
stale scenes. Actuator (motor) control latency also can be a
factor, because moving the steering motor to a desired position takes
time. 
In our platform, the combined latency is measured to be around
50 ms. In other words, in the recorded video and control data, a
control action (left, center, right) taken by the CNN appears to be
applied 50 ms later. Considering that camera's framerate is 30Hz
(33.$\overline{\mbox{33}}$ ms/frame), this is about two frames after the 
control action.
%% We experimentally measured the camera
%% latency and found it to be around 50-100 ms.
If this value is too high, control performance may suffer.
Our initial prototype suffered this problem as the observed camera and
control latency was as high as 300 ms, which negatively affected
control performance.
For reference, the latency of human perception is known to be as fast
as 13 ms~\cite{ThomasBurger2015}. 
% https://www.pubnub.com/blog/2015-02-09-how-fast-is-realtime-human-perception-and-technology/

Despite these issues, the trained models were still able to
achieve a reasonable degree of accuracy, successfully navigating
several different tracks we used for training. In our testing, we found 
that the DeepPicar could remain on a fairly challenging track for over 
10 minutes at a moderate speed (50\% throttle), at which point we stopped
the experiment, and more than one minute at a higher speed (75\%
throttle)~\footnote{A collection of self-driving videos can be found
  at: \url{https://photos.app.goo.gl/ce93sU7jPk4ywO8u2}}. Running at
higher speed is inherently more challenging because the CNN controller
has less time to recover from mistakes (bad predictions).  Also, we
find that the prediction accuracy is significantly affected by the
quality of training data as well as various environmental factors such
as lighting conditions. We plan to investigate more systematic ways
to improve the CNN's prediction accuracies.
